Titeln på avhandlingens språk

### 2022-12-02 body/part1/background.tex
working

\chapter{Introduction}

\section{Background \& Related Work}

@@

\subsection{Theoretical background}
\textcite{Satya2009Case}.
Other papers detailing the advantages of Edge computing.

@@

\section{Scope of this thesis}

What research questions are being tackled?

How will human-in-the-loop/\ac{WCA} systems look in production?

How will they look in large scale deployments?

How many applications can given systems support?

How can these dimensions be studied?
Here's where modeling and empirical data-driven research comes in.

Is there potential for optimization?
Yes, trading of system responsiveness for decrease in system load.
Underlying principle under aperiodic, adaptive sampling.
The potential for optimization is novel and stems from \textcite{Olguin:ImpactWCA2021}.

>>>>

\chapter{Background \& Related Work}

@@

\section{Theoretical background}
\textcite{Satya2009Case}.
Other papers detailing the advantages of Edge computing.

@@

### 2022-12-02 body/part1/introduction.tex
struct
working

\chapter{Introduction}

\section{Background \& Related Work}

Deep focus on theoretical background.

\subsection{Theoretical background}
\textcite{Satya2009Case}.
Other papers detailing the advantages of Edge computing.

Next, what applications benefit from the Edge?
\ac{WCA}! Mention both the main body of work (by Satya et.\ al.) but also other authors out there.
Main points: prototypes and characterization of system latencies and load.

\section{Presentation of research questions}

How will human-in-the-loop/\ac{WCA} systems look in production?

How will they look in large scale deployments?

How many applications can given systems support?

How can these dimensions be studied?
Here's where modeling and empirical data-driven research comes in.

Is there potential for optimization?
Yes, trading of system responsiveness for decrease in system load.
Underlying principle under aperiodic, adaptive sampling.
The potential for optimization is novel and stems from \textcite{Olguin:ImpactWCA2021}.

>>>>

\chapter{Introduction}\label{chap:introduction}

\todo[inline]{
    Introduction here. Why do we need this work?
}

\todo[inline]{
    Reorganize EdgeDroid, Cleave, Ainur, together.

    Pull in multi-loop results for EdgeDroid?

    Think about experiments.
}

\section{Summary of the key contributions of this thesis}

This thesis presents three core contributions to the existing body of research in edge computing, as well as a number of secondary ones.
Firstly, we introduce a methodological approach to studying system responsiveness versus resource consumption trade-offs in edge-bound feedback-loop systems such as \ac{WCA} and \ac{NCS}.
This approach is based on the emulation of the client-side workload, while maintaining the \emph{real} server-side process as well as network stack.
We validate this methodology by example, presenting implementations of its application in the context of \ac{WCA} and \acp{NCS}, and introduce a software framework for the orchestration of the edge computing testbeds necessary for its implementation.

Secondly, we present a deeper exploration of this methodology for \ac{WCA} and introduce the first ever model for the end-to-end emulation of human timing behavior in \ac{WCA}.
This model is built from a thorough characterization of these behaviors, the data for which we obtain from a comprehensive human-subject study.

Finally, as an extension and application of our first and second contributions, we explore the implications of our methodology and human user model on the optimization potential of \ac{WCA} deployments.
In particular, we study the sampling and energy consumption behaviors of these systems with and without considering realistic human behavior.
We conclude that significant improvements can be achieved through the use of our human user model.

\subsection{Overview of each included paper}

\cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid,paper:olguinmunoz2022cleave,paper:olguinmunoz2022ainur} introduce and discuss this methodology and the necessary tools for its implementation.
\cref{paper:olguinmunoz2018demoscaling} presents a short, high level overview of this methodological approach as applied to \ac{WCA}, making very straightforward assumptions about human behavior in \ac{WCA} applications.
\cref{paper:olguinmunoz2019edgedroid} extends the discussion from \cref{paper:olguinmunoz2018demoscaling}, providing more details on the implementation of the necessary measurement framework and the tooling employed.
This work presents the first empirical results obtained with our methodological approach, providing thus an initial validation of its utility for research.
\cref{paper:olguinmunoz2019edgedroid} also discusses in detail the assumptions that were made about human behavior and reactions.
We assume in these works a human which is impervious to poor system performance, and suffers no annoyance, fatigue, frustration, nausea or other shortcomings of real human users.
The result is a model of a user who responds in a precisely reproducible and deterministic manner to the same system stimulus every time.

\cref{paper:olguinmunoz2022cleave}, on the other hand, presents a discussion on the implementation of this methodology for \acp{NCS}.
As discussed above, these applications are similarly difficult as \ac{WCA} to benchmark (in particular at scale on multi-tenant systems) due to their client-side complexity, but in particular because of their often extreme sensitivity to latency.
\todo[inline]{More argumentation or refer to previous section.}
We apply thus in this work our methodology to \ac{NCS} through the implementation of a tool for the emulation and subsequent deployment of these systems on edge computing infrastructure.
This tool allows us to both emulate the physical components of a relatively simple control system plant and deploy real algorithms for its control.
The software acts as a middleware which abstracts away the network from the development of these workloads, allowing for quick prototyping and deployment.
This work also showcases the scalability and flexibility of our approach, allowing us to deploy scenarios with a large number of loops without the need for domain-specific hardware.

\cref{paper:olguinmunoz2022ainur} presents a tangential contribution.
It introduces the software framework used in \cref{paper:olguinmunoz2023realistic,paper:olguinmunoz2022cleave} for the orchestration of the edge computing testbeds on which the developed tools were deployed.
This framework represents an ancillary contribution, crucial for the research presented in these works.
Without it, the experimental approach described in these works would not have been feasible.

Next, in \cref{paper:olguinmunoz2021impact,paper:olguinmunoz2023realistic} we delve deep into the implementation of our methodology for \ac{WCA}, as well as study its implications for optimization of these systems.
Although the above approach to human behavior in \ac{WCA} discussed represents a useful initial approximation, it is nonetheless not a realistic model of it.
In \cref{paper:olguinmunoz2021impact} we therefore take the first major step towards such a model, presenting a deep characterization of human behavior in these applications.
We develop this characterization through a human subject study with a cohort of \num{40} participants who were asked to interact with an instrumented \ac{WCA} application.
System responsiveness is altered in real-time during each execution of the task, and we record participants' reactions by measuring task- and system-related metrics, as well as biometrics from sensors placed on their bodies.
Participants were also asked to fill out two personality indicator questionnaires, allowing us to later correlate individual personality traits and measured reaction to changes in system responsiveness.

\cref{paper:olguinmunoz2023realistic} then concludes this line of work, building upon the insights and data obtained in \cref{paper:olguinmunoz2021impact} to develop the first ever data-driven model of human timings for \ac{WCA}.
The model is validated against previously obtained results, both through simulated, controlled executions and deployments on a real edge computing testbed.
This work also explores potential implications of this model for \ac{WCA} system optimization potential, particularly in the domains of energy consumption and sampling strategies.

\section{Structure of this thesis}

\todo[inline]{This thesis is structured as follows.}

### 2022-12-02 body/part1/overview.tex
working

\chapter{Overview of this thesis' contributions}

\todo[inline]{
    Methodology: a system of methods used in a particular area of study or activity.
}

\section{Scope of this thesis}

We study in this thesis a methodological approach to studying system responsiveness versus resource consumption trade-offs in edge-bound feedback-loop systems such as \ac{WCA} and \ac{NCS}.
This approach is based on the emulation of the client-side workload, while maintaining the \emph{real} server-side process as well as network stack.
We validate this methodology by example, presenting a deep exploration of its implementation and applications in the context of \acl{WCA}, and a initial, more shallow foray into its applications in the context of \ac{NCS}.

More specifically, in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} we introduce this methodology for \ac{WCA}.
\cref{paper:olguinmunoz2018demoscaling} presents a short, high level overview of this methodological approach, making very straightforward assumptions about human behavior in \ac{WCA} applications.
\cref{paper:olguinmunoz2019edgedroid} extends the discussion from \cref{paper:olguinmunoz2018demoscaling}, providing more details on the implementation of the necessary measurement framework and the tooling employed.
This work presents the first empirical results obtained with our methodological approach, providing thus an initial validation of its utility for research.
Finally, \cref{paper:olguinmunoz2019edgedroid} discusses in detail the assumptions that were made about human behavior and reactions.
We assume in these works a human which is impervious to poor system performance, and suffers no annoyance, fatigue, frustration, nausea or other shortcomings of real human users.
The result is a model of a user who responds in a precisely reproducible and deterministic manner to the same system stimulus every time.

Although the above approach to human behavior in \ac{WCA} represents a useful initial approximation, it is nonetheless not a realistic model of it.
In \cref{paper:olguinmunoz2021impact} we therefore take the first major step towards such a model, presenting a deep characterization of human behavior in these applications.
We develop this characterization through a human subject study with a cohort of \num{40} participants who were asked to interact with an instrumented \ac{WCA} application.
System responsiveness is altered in real-time during each execution of the task, and we record participants' reactions by measuring task- and system-related metrics, as well as biometrics from sensors placed on their bodies.
Participants were also asked to fill out two personality indicator questionnaires, allowing us to later correlate individual personality traits and measured reaction to changes in system responsiveness.

\cref{paper:olguinmunoz2023realistic} then concludes this line of work, building upon the insights and data obtained in \cref{paper:olguinmunoz2021impact} to develop the first ever data-driven model of human timings for \ac{WCA}.
The model is validated against previously obtained results, both through simulated, controlled executions and deployments on a real edge computing testbed.
This work also explores potential implications of this model for \ac{WCA} system optimization potential, particularly in the domains of energy consumption and sampling strategies.

On the other hand, \cref{paper:olguinmunoz2022cleave} presents an extension to the methodology introduced and developed in \cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} to the field of \aclp{NCS}.
We implement a tool for the emulation of these systems on edge computing infrastructure, and deploy it on an edge computing testbed for validation.
In particular, this work showcases the scalability and flexibility of our approach, allowing us to deploy scenarios with a large number of loops without the need for domain-specific hardware.

Finally, \cref{paper:olguinmunoz2022ainur} presents the software framework used in \cref{paper:olguinmunoz2023realistic,paper:olguinmunoz2022cleave} for the orchestration of the edge computing testbeds on which the developed tools were deployed.
This framework represents an ancillary contribution, tangential to yet crucial for the research presented in these works.
Without it, the experimental approach described in these works would not have been feasible.

In the following sections, we will summarize in detail the key contributions of each of the above-mentioned works.
\cref{paper:olguinmunoz2018demoscaling,paper:olguinmunoz2019edgedroid} are discussed in detail in \cref{summary:2018demo,summary:2019edgedroid}.
Next, in \cref{summary:model} we present \cref{paper:olguinmunoz2021impact} (\cref{summary:2021impact}) and \cref{paper:olguinmunoz2023realistic} (\cref{summary:2023realistic}).
\cref{paper:olguinmunoz2022cleave} is summarized in \cref{summary:2022cleave}, and we finalize this chapter with a summary of \cref{paper:olguinmunoz2022ainur} in \cref{summary:2022ainur}.

>>>>

\chapter{Key contributions and results}\label{chap:contributions}

\todo[inline]{Reorganize}
